---
title: "1.1 Template"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
editor_options: 
  markdown: 
    wrap: 80
---

```{r setup, include=FALSE}
library(dplyr)
hdr_path <- case_when(
  file.exists(here::here("R", "header-include.R")) ~ here::here("R", "header-include.R"), 
  file.exists(here::here("tutorials", "R", "header-include.R")) ~ here::here("tutorials", "R", "header-include.R")
  )
  

source(hdr_path)
```


## Welcome

Hello, and welcome to **Introduction to data**!


```{r quiz-rows-columns}
quiz(
  caption = "",
  question("how many observations and how many variables are in the `hsb2` dataset? ",
    allow_retry = TRUE, random_answer_order = TRUE,
    answer("200 variables", message = "Nope, try again!"),
    answer("200 observations", correct = TRUE),
    answer("11 variables", correct = TRUE),
    answer("11 observations", message = "Nope, try again!")
  ), 
  question("What do you think: Do cars with bigger engines use more fuel than cars with smaller engines?",
    allow_retry = TRUE, 
    random_answer_order = TRUE, 
    type = "learnr_radio",
    answer("Cars with bigger engines use _more_ fuel.", correct = TRUE, message = "You guessed that there is a positive relationship between engine size and fuel efficiency. Now let's test your hypothesis against data."),
    answer("Cars with bigger engines use _less_ fuel.", correct = TRUE, message = "You guessed that there is a negative relationship between engine size and fuel efficiency. Now let's test your hypothesis with data.")
    )
  )




```

## What's next?

`r emo::ji("one")` [Lesson 1.1: Language of data](https://oferengel-posit.shinyapps.io/01-01-lesson/)

`r emo::ji("two")` [Lesson 1.2: Types of studies](https://oferengel-posit.shinyapps.io/01-02-lesson/)

`r emo::ji("three")` [Lesson 1.3: Case study](https://oferengel-posit.shinyapps.io/01-03-lesson/)

## Hash and submit

```{r encoder, echo=FALSE, context="server"}
encoder_logic()
```

```{r encode, echo=FALSE}
learnrhash::encoder_ui(ui_before = hash_encoder_ui)
```

### Submit your hash in a form

After creating the hash, please copy it, navigate to [this form](`r submit_hash_url`) and fill it in and submit. You will need to submit the form once for every lesson you complete in this tutorial.

```{r img-submit, message=FALSE, warning=FALSE, fig.align='center', out.width="50%"}




p_img <- ifelse(
  file.exists(
    here("tutorials", "images", "Submit-hash.jpg")
    ),
  here("tutorials", "images", "Submit-hash.jpg"),
  here("images", "Submit-hash.jpg")
  )

p_img
if (file.exists(p_img))
  knitr::include_graphics(p_img)
```
