---
title: R Basics
author: Your name here
date: The date here
format: html
editor: source
editor_options: 
  chunk_output_type: console
---


this is my qmd file, it consists of texts for humans like this, and it consists of code see below.



```{r}
#| label: setup
#| include: false

# Remember that all tidyverse packages have already been installed for you, 
# but you will need to *load* the ones you want to use.
library(tidyverse)

# This code loads the data-set and stores it in an object
arbuthnot <- read_csv(here::here("labs/shared/data/arbuthnot.csv"))  # Ctrl + Enter

```

## R Basics

In this milestone, we will work with the `arbuthnot` data set. The Arbuthnot data set refers to the work of [Dr. John Arbuthnot](https://en.wikipedia.org/wiki/John_Arbuthnot), an 18<sup>th</sup> century physician, writer, and mathematician. The data records the annual counts of male and female births in **London** from 1629 to 1710. It is considered one of the earliest examples of using data for statistical analysis and hypothesis testing. 

Before continuing, try to guess: will there be more baby girls or baby boys born in any one year? 

We can view the data by running the code below or by typing the name of the dataset into the console. Be careful the spelling and capitalization you use! `R` is case sensitive, so if you accidentally type `Arbuthnot` `R` will tell you that object cannot be found.


## Recreation

The data we will use in our investigation is saved as an R object named `arbuthnot`. Use R to inspect the contents of `arbuthnot`. We will look at two types of functions: 

1. Functions that use the entire dataset as an input (argument). Those are functions such as `summary(___)`, `nrow(___)` and others. Replace the `___` with the dataset object. In our case, that would be `arbuthnot`) 
2. Functions that use one of the columns of the dataset as input (argument). Those are functions such as `mean(___)`, `min(___)`, `max(___)`. Replace the `___` with a column from our dataset. For example, if you want to find the largest year in the `$year` column, you will need to type `max(arbuthnot$year)`. 


Functions that use the entire dataset as an argument: 

- Display the contents of `arbuthnot` dataset
- Use `nrow()` to find the number of observations in the dataset
- Use `names()` to find the variables in the dataset (column names)
- Use `summary()` to generate a summary of the data set. 
- Use `mean()` to find the average number of boys' and girls' birth count across all years
- Use `median()` to find the median number of boys' and girls' birth count across all years
- Use `max()` and `min()` to find the maximum and minimum number of boys' and girls' birth count 
- Use `quantile()` to find the lower and upper quartile of boys' and girls' 
- Look up the help page and for `fivenum()` and use it to find Tukey's five number summary for boys' and girls'. Compare it to the results from above!

*Hint:* you may want to look up the help pages for each of these functions.

Write your code in the following chunk:


```{r}
#| label: recreation




```



## Extension part 1

Explore `arbuthnot` in a new way. Consider using a new function, or something that you can learn from a help page. Alternatively, try to learn more about the other variables in `covid`. Keep in mind that we will study how to make plots for the next milestone.

**For your first extension, we encourage you to think small.** Throughout this course, you will build many skills to explore data; for this first week, your work will likely be fairly straightforward. Consider _limiting yourself to 30-45 minutes_ for this portion of your work; at the end of that time interval, stop working and make notes on your progress, dead-ends, and aspirations -- bringing those to your group for discussion during the milestone session.


Some questions you might want to explore (but feel free to ignore them and explore your own questions if you like!): 
1. Functions such as `mean()`, `median()` and `mode()` are known as measures of central tendency. Which group has a larger measure of central tendency: boys or girls? 
2. Function such as `range()`, `var()`, `sd()`, `range()`, `IQR()` are known as measures of dispersion. Which group has a larger measure of dispersion: boys or girls?
3. Is the distribution of births right skewed or left skewed? How can we tell? What does this mean?
4. Which year(s). have the lowest birthrate? The highest birthrate? 
5. Does the year that have the highest (lowest) birthrate for males also have the highest (lowest) for females?

Write your extension code in the following chunk:


```{r}
#| label: extension-1



```


Some additional information about the `arbuthnot` dataset:

Arbuthnot used the data to demonstrate a consistent excess of male births over female births. He argued this supported a divine plan for population balance, given that men were more likely to die earlier due to war, dangerous work, etc. There were several historical events that may have influenced the birth rate over this time: 

1. London experienced major population disruptions, such as the Great Plague of 1665. 
2. The English Civil War (1642–1651) and other social disruptions may have reduced fertility rates or registration completeness.



## Extension part II

In the previous sections you explored Arbuthnot's baptism data. In this section you are invited to repeat these steps, but for present day birth records in the United States. The data are stored in a csv file within the `data` directory. Load the data frame in a similar way to the way we've loaded the `arbuthnot` dataset, and explore the data in a similar way. Comment on similarities and differences when comparing the `arbuthnot` dataset to the `present` dataset! 


```{r}
#| label: extension-2



```

